Examining spatial equity and accessibility to a public bicycle share program using a balanced floating catchment area approach

### 2021-03-13 Bike-share-spatial-equity/Bike-share-spatial-equity.Rmd
Major reorganization of repository to clean up files and to create (mostly data) package.

```{r, include=FALSE}
rm(list = ls())
```

```{r, include=FALSE}
library(data.table)
library(dplyr)
library(gdistance)
library(kableExtra)
library(raster)
library(rgdal)
library(sf)
library(tidyverse)
library(units)
library(disk.frame)
library(pycno)
library(readr)
library(r5r) # the r5r package requires Java Development Kit version 11, which can be downloaded from https://www.oracle.com/java/technologies/javase-jdk11-downloads.html
library(tinytex)
library(ggplot2)
theme_set(theme_bw())
```

```{r, include=FALSE}
equity_stations <- data.frame(longitude = c(	
-79.84364993846138, -79.8362839215856, -79.83144521746212, -79.82217214949738, -79.8160211746482, -79.8192304372373, -79.82500255113169, -79.8268599805827, -79.83888834759689, -79.82740849222654, -79.82342541216097, -79.84240606419081), latitude = c(43.265340158000654, 43.25632467284136, 43.25528943845642, 43.25291050074978, 43.25147804403345, 43.24444659758296, 43.2421627336722, 43.24634382671116, 43.249432741404554, 43.251126444296524, 43.24892849896579, 43.25831920064561))
```

@@

```{r, include=FALSE}
load("hamilton_cma.RData")
load("hamilton_da_2016.RData")
load("population_sobi.RData") #interpolated population in SoBi area
load("sobi_hubs.RData") #location of SoBi hubs, minus Van Wagner's beach
load("sobi_hubs_original.RData") #SoBi hubs, minus ERI hubs
load("sobi_service.RData") #core service area for SoBi
load("ttm_walk.RData") #travel time matrix for interpolated population in SoBi area
load("ttm_walk_original.RData") #travel time matrix for system without ERI hubs
load("accessibility.RData") #full accessibility table for SoBi system
load("accessibility_map_original.RData") #map of accessibility in original SoBi system
load("accessibility_map_equity.RData") #map of accessibility in current system with ERI
load("hamilton_da_sobi.RData")
load("population_sobi_original.RData") #population in original system
load("mhi.RData") #median household income by DA
```

@@

```{r, include=FALSE}
sobi_service <- sobi_service %>%
  dplyr::filter(AREA_NAME == "Core Service Area")
```

```{r, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "The core service area of SoBi Hamilton is outlined in blue. Dissemination areas in the Hamilton Census Metropolitan Area (CMA) are outlined in grey."}
ggplot() + 
  geom_sf(data = hamilton_da_2016,
          fill = NA,
          color = "gray") + 
  geom_sf(data = sobi_service,
          fill = NA,
          colour = "blue")
```

```{r, include=FALSE}
eri_hubs <- sobi_hubs %>% 
             filter(OBJECTID>=117) %>% filter(OBJECTID<=128)
```

```{r figure-1, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "The spatial distribution of PBSP stations in Hamilton, Ontario. The service area of the PBSP is outlined in blue, and dissemination areas within the Hamilton CMA that can reach a SoBi hub within 30 minutes of walking are outlined in grey."}
ggplot() + 
  geom_sf(fill = "antiquewhite1") +
  geom_sf(data = hamilton_da_sobi, aes(color = "gray")) +
  geom_sf(data = sobi_service, fill = NA, aes(color = "blue")) + 
  geom_sf(data = sobi_hubs_original, size = 1, aes(color = "red")) +
  geom_sf(data = eri_hubs, size = 1, aes(color = "green"))+
  scale_color_identity(name = "Station Type",
                       breaks = c("red", "green"),
                       labels = c("Original Station", "Equity Station"),
                       guide = "legend")
```

@@

The first step is to allocate the population to be serviced by each PBSP station: 
$$P_j = {\sum_{i = 1}^{n} P_i{w_{ij}}}$$
In the second step, the level of service at each bicycle share station is the capacity at the location (i.e., the maximum number of bicycle racks) divided by its estimated service population:
$$L_j = \frac {S_j}{P_j} = \frac {S_j}{{\sum_{i = 1}^{n} P_i{w_{ij}}}}$$

Finally, accessibility is calculated:
$$A_i = {\sum_{j = 1}^{J} L_j{w_{ji}}} = {\sum_{j = 1}^{J} \frac {S_j{w_{ji}}}{\sum_{i = 1}^{n} P_i{w_{ij}}}}$$

Weights are then calculated:
$$w_{ij} = f({c_{ij}})$$

The approach replaces the weights in the calculations above, and uses instead a set of suitably normalized weights as follows:
$${w_{ij}^{i} = \frac {w_{ij}}{\sum_{j = 1}^{J} {w_{ji}}}}$$
and
$${w_{ij}^{j} = \frac {w_{ij}}{\sum_{j = 1}^{J} {w_{ji}}}}$$

These weights satisfy the following properties:
$$\sum_{j = 1}^{J} {w^i_{ji}} = 1$$

and
$$\sum_{j = 1}^{J} {w^j_{ji}} = 1$$

Finally, accessibility is calculated:
$$A_i = {\sum_{j = 1}^{J} \frac {S_j{w^j_{ij}}}{\sum_{i = 1}^{n} P_i{w^i_{ij}}}}$$

@@

```{r figure-2, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "Population in all dissemination areas in the SoBi Hamilton core service area and within 30 minutes of walking to a bike share station."}
ggplot() +
  geom_sf(data = hamilton_da_sobi,
          aes(fill = POP2011),
          color = NA) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_fill_distiller(palette = "YlOrRd", direction = 1)
```

```{r figure-4, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "Interpolated population cells in the SoBi Hamilton core service area (outlined in black) and within 30 minutes walk to the core service area. Geographic features have been extracted. Dissemination area polygons are outlined in grey."}
ggplot() + 
  geom_sf(data = population_sobi,
          aes(color = population)) +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1) +
  geom_sf(data = hamilton_da_sobi,
          fill = NA,
          color = "lightgray") + 
  geom_sf(data = sobi_service,
          fill = NA,
          colour = "blue")
```

```{r figure-5, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "Interpolated population units in the SoBi Hamilton core service area (outlined in black) and within 30 minutes walk to the core service area. Green circles represent the location of the equity stations that were added in 2018. Red circles represent the original stations. The service area of the PBSP is outlined in blue, and dissemination areas from the 2016 Canadian Census are outlined in light grey."}
ggplot() + 
  geom_sf(fill = "antiquewhite1") +
  geom_sf(data = population_sobi, aes(color = population)) +
  scale_color_distiller(palette = "YlOrRd", direction = 1) +
  geom_sf(data = hamilton_da_sobi, fill = NA, color = "lightgray") + 
  geom_sf(data = sobi_hubs_original, size = 1.5, shape = 21, fill = "red") +
  geom_sf(data = eri_hubs, size = 1.5, shape = 21, fill = "green") +
  geom_sf(data = sobi_service, fill = NA, colour = "blue")
```

@@

```{r echo=FALSE}
# current system with equity stations
ttm_walk <- left_join(ttm_walk, 
                      population_sobi,
                      by = "UID")
```

```{r echo=FALSE} 
# original system without equity stations
ttm_walk_o <- left_join(ttm_walk_o, 
                      population_sobi_original,
                      by = "UID")
```

```{r, echo=FALSE}
ttm_walk_o <- ttm_walk_o[-c(5,7:8)]
```

```{r echo=FALSE}
ttm_walk <- ttm_walk %>% dplyr::select(-geometry)

ttm_walk_o <- ttm_walk_o %>% dplyr::select(-geometry)
```

```{r, include=FALSE}
# current system with equity stations
ttm_walk <- left_join(ttm_walk, sobi_hubs %>% dplyr::select(OBJECTID, RACKS_AMOU), by=c("OBJECTID" = "OBJECTID"))
ttm_walk <- ttm_walk %>% dplyr::select(-geometry)
head(ttm_walk)
```

```{r, include=FALSE}
# original system without equity stations
ttm_walk_o <- left_join(ttm_walk_o, sobi_hubs_original %>% dplyr::select(OBJECTID, RACKS_AMOU), by=c("OBJECTID" = "OBJECTID"))
ttm_walk_o <- ttm_walk_o %>% dplyr::select(-geometry)
head(ttm_walk_o)
```

```{r echo=FALSE}
#data frames for current and original systems
names(ttm_walk) <- c('UID', 'hub', 'travel_time', 'population', 'racks')

names(ttm_walk_o) <- c('UID', 'hub', 'travel_time', 'population', 'racks')
```

### Threshold of 3 Minutes Walk

```{r, echo=FALSE}
#minimum threshold for accessibility calculations
dist_threshold <- 3
```

```{r echo=FALSE}
ttm_walk_min <- ttm_walk %>%
  mutate(impedance_binary = ifelse(travel_time <= dist_threshold, 1, 0))

ttm_walk_o_min <- ttm_walk_o %>%
  mutate(impedance_binary = ifelse(travel_time <= dist_threshold, 1, 0))
```

```{r echo=FALSE}
sum_b1_min <- ttm_walk_min %>%
  group_by(UID) %>%
  summarize(sum_b1_min = sum(impedance_binary),
            .groups = "drop")

sum_b1_o_min <- ttm_walk_o_min %>%
  group_by(UID) %>%
  summarize(sum_b1_o_min = sum(impedance_binary),
            .groups = "drop")
```

```{r echo=FALSE}
sum_b2_min <- ttm_walk_min %>%
  group_by(hub) %>%
  summarize(sum_b2_min = sum(impedance_binary))

sum_b2_o_min <- ttm_walk_o_min %>%
  group_by(hub) %>%
  summarize(sum_b2_o_min = sum(impedance_binary))
```

```{r echo=FALSE}
ttm_walk_min <- ttm_walk_min %>%
  left_join(sum_b1_min, by = "UID") %>%
  left_join(sum_b2_min, by = "hub")

ttm_walk_o_min <- ttm_walk_o_min %>%
  left_join(sum_b1_o_min, by = "UID") %>%
  left_join(sum_b2_o_min, by = "hub")
```

```{r echo=FALSE}
ttm_walk_min <- ttm_walk_min %>%
  dplyr::filter(sum_b1_min > 0)

ttm_walk_o_min <- ttm_walk_o_min %>%
  dplyr::filter(sum_b1_o_min > 0)
```

```{r echo=FALSE}
ttm_walk_min <- ttm_walk_min %>% 
  dplyr::mutate(balanced_impedance_1 = impedance_binary/sum_b1_min, 
         balanced_impedance_2 = impedance_binary/sum_b2_min)

ttm_walk_o_min <- ttm_walk_o_min %>% 
  dplyr::mutate(balanced_impedance_1 = impedance_binary/sum_b1_o_min, 
         balanced_impedance_2 = impedance_binary/sum_b2_o_min)
```

```{r echo=FALSE}
ttm_walk_min %>% 
  dplyr::filter(hub == "46") %>%
  pull(balanced_impedance_2) %>%
  sum()

ttm_walk_o_min %>% 
  dplyr::filter(hub == "46") %>%
  pull(balanced_impedance_2) %>%
  sum()
```

```{r, include=FALSE}
sobi_los_min <- ttm_walk_min %>%
  dplyr::filter(impedance_binary > 0) %>% # filter impedance == 0 to avoid divisions by zero
  group_by(hub) %>%
  summarize(sobi_los_min = first(racks) / sum((population * balanced_impedance_1)),
            .groups = "drop")

sobi_los_o_min <- ttm_walk_o_min %>%
  dplyr::filter(impedance_binary > 0) %>% # filter impedance == 0 to avoid divisions by zero
  group_by(hub) %>%
  summarize(sobi_los_o_min = first(racks) / sum((population * balanced_impedance_1)),
            .groups = "drop")
```

```{r, include=FALSE}
ttm_walk_min <- ttm_walk_min %>%
  left_join(sobi_los_min, by = "hub")

ttm_walk_o_min <- ttm_walk_o_min %>%
  left_join(sobi_los_o_min, by = "hub")
```

```{r, include=FALSE}
sobi_accessibility_min <- ttm_walk_min %>%
  group_by(UID) %>%
  summarize(accessibility_min = sum(sobi_los_min * balanced_impedance_2),
            .groups = "drop")

sobi_accessibility_o_min <- ttm_walk_o_min %>%
  group_by(UID) %>%
  summarize(accessibility_o_min = sum(sobi_los_o_min * balanced_impedance_2),
            .groups = "drop")
```

```{r, echo=FALSE}
#current system with equity stations
sobi_accessibility_min %>% 
  pull(accessibility_min) %>% 
  sum()

sobi_los_min %>% 
  pull(sobi_los_min) %>%
  sum()
```

```{r, echo=FALSE}
#original system without equity stations
sobi_accessibility_o_min %>% 
  pull(accessibility_o_min) %>% 
  sum()

sobi_los_o_min %>% 
  pull(sobi_los_o_min) %>%
  sum()
```

```{r, include=FALSE}
population_sobi <- population_sobi %>%
  left_join(sobi_accessibility_min,
            by = "UID")

population_sobi_original <- population_sobi_original %>%
  left_join(sobi_accessibility_o_min,
            by = "UID")
```

```{r, echo=FALSE}
#accessibility plot for current system with equity stations at 3 minutes walk
ggplot(data = population_sobi,
       fill = accessibility_min) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1)
```

```{r, echo=FALSE}
#accessibility plot for original system without equity stations at 3 mins walk
ggplot(data = population_sobi_original) +
  geom_sf(aes(color = accessibility_o_min)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1)
```

```{r, echo=FALSE}
#level of equitable service in current system at 3 mins walk
population_sobi %>%
  mutate(Lei_min = accessibility_min - sum(sobi_los_min$sobi_los_min)/nrow(population_sobi %>% drop_na(accessibility_min))) %>%
  ggplot() +
  geom_sf(aes(color = Lei_min)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_color_gradient2()
```

@@

```{r, echo=FALSE}
ggplot() +
  geom_sf(data = population_sobi_original,
          aes(color = accessibility_o_min)) +
  geom_sf(data = hamilton_da_sobi,
          fill = NA) + 
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1)
```

@@

### Threshold of 10 Minutes Walk

```{r, echo=FALSE}
dist_threshold <- 10
```

@@

```{r, echo=FALSE}
#accessibility of current system at 10 mins walk
 cowplot::ggdraw() +
 cowplot::draw_plot(map_equity, 0, 0, 1, 1) +
 cowplot::draw_plot(legend, 0.05, .6, 0.2, 0.2)
```

>>>>

```{r load-packages, include=FALSE}
library(data.table)
library(dplyr)
library(gdistance)
library(kableExtra)
library(raster)
library(rgdal)
library(sf)
library(sobiEquity)
library(tidyverse)
library(units)
#library(disk.frame)
#library(pycno)
library(readr)
#library(r5r) # the r5r package requires Java Development Kit version 11, which can be downloaded from https://www.oracle.com/java/technologies/javase-jdk11-downloads.html
library(tinytex)
library(ggplot2)
theme_set(theme_bw())
```

```{r}
data("hamilton_cma")
data("hamilton_da_2016")
data("population_50x50")
data("sobi_service")
data("sobi_hubs")
data("ttm_walk")
```

```{r coordinates-of-equity-stations, include=FALSE}
# define the coordinates of the equity stations
# equity_stations <- data.frame(longitude = c(-79.84364993846138, 
#                                             -79.8362839215856,
#                                             -79.83144521746212, 
#                                             -79.82217214949738, 
#                                             -79.8160211746482, 
#                                             -79.8192304372373, 
#                                             -79.82500255113169, 
#                                             -79.8268599805827, 
#                                             -79.83888834759689, 
#                                             -79.82740849222654, 
#                                             -79.82342541216097, 
#                                             -79.84240606419081), 
#                               latitude = c(43.265340158000654, 
#                                            43.25632467284136, 
#                                            43.25528943845642, 
#                                            43.25291050074978, 
#                                            43.25147804403345, 
#                                            43.24444659758296, 
#                                            43.2421627336722, 
#                                            43.24634382671116,
#                                            43.249432741404554,
#                                            43.251126444296524, 
#                                            43.24892849896579, 
#                                            43.25831920064561))
```

@@

```{r load-data, include=FALSE}
#load("accessibility.RData") #full accessibility table for SoBi system
#load("accessibility_map_original.RData") #map of accessibility in original SoBi system
#load("accessibility_map_equity.RData") #map of accessibility in current system with ERI
# load("hamilton_cma.RData")
# load("hamilton_da_2016.RData")
# load("hamilton_da_sobi.RData")
# load("mhi.RData") #median household income by DA
# load("population_sobi.RData") #interpolated population in SoBi area
#load("population_sobi_original.RData") #population in original system
# load("sobi_hubs.RData") #location of SoBi hubs, minus Van Wagner's beach
# load("sobi_hubs_original.RData") #SoBi hubs, minus ERI hubs
# load("sobi_service.RData") #core service area for SoBi
# load("ttm_walk.RData") #travel time matrix for interpolated population in SoBi area
# load("ttm_walk_original.RData") #travel time matrix for system without ERI hubs
```

@@

```{r hamilton-and-sobi-service-area, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "The core service area of SoBi Hamilton is outlined in blue. Dissemination areas in the Hamilton Census Metropolitan Area (CMA) are outlined in grey."}
ggplot() + 
  geom_sf(data = hamilton_da_2016,
          fill = NA,
          color = "gray") + 
  geom_sf(data = sobi_service,
          fill = NA,
          colour = "blue")
```

```{r filter-eri-hubs, include=FALSE}
# eri_hubs <- sobi_hubs %>% 
#              filter(OBJECTID>=117 & OBJECTID<=128)
```

```{r sobi-stations-in-hamilton, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "The spatial distribution of PBSP stations in Hamilton, Ontario. The service area of the PBSP is outlined in blue, and dissemination areas within the Hamilton CMA that touch the bounding box of the PBSP's service area are outlined in grey."}
ggplot() + 
  #geom_sf(fill = "antiquewhite1") +
  geom_sf(data = hamilton_da_2016 %>% 
            dplyr::filter(sobi_service == "Yes"),
          color = "gray") +
  geom_sf(data = sobi_service, 
          fill = NA,
          color = "blue") + 
  geom_sf(data = sobi_hubs %>% 
            dplyr::filter(hub_type == "Conventional"), 
          size = 1.5, 
          shape = 21, 
          fill = "red") +
  geom_sf(data =  sobi_hubs %>% 
            dplyr::filter(hub_type == "ERI"), 
          size = 1.5, shape = 21, 
          fill = "green")
```

@@

The first step is to allocate the population to be serviced by each PBSP station: 
$$
P_j = {\sum_{i = 1}^{n} P_i{w_{ij}}}
$$
In the second step, the level of service at each bicycle share station is the capacity at the location (i.e., the maximum number of bicycle racks) divided by its estimated service population:
$$
L_j = \frac {S_j}{P_j} = \frac {S_j}{{\sum_{i = 1}^{n} P_i{w_{ij}}}}
$$

Finally, accessibility is calculated:
$$
A_i = {\sum_{j = 1}^{J} L_j{w_{ji}}} = {\sum_{j = 1}^{J} \frac {S_j{w_{ji}}}{\sum_{i = 1}^{n} P_i{w_{ij}}}}
$$

Weights are then calculated:
$$
w_{ij} = f({c_{ij}})
$$

The approach replaces the weights in the calculations above, and uses instead a set of suitably normalized weights as follows:
$$
{w_{ij}^{i} = \frac {w_{ij}}{\sum_{j = 1}^{J} {w_{ji}}}}
$$
and
$$
{w_{ij}^{j} = \frac {w_{ij}}{\sum_{j = 1}^{J} {w_{ji}}}}
$$

These weights satisfy the following properties:
$$
\sum_{j = 1}^{J} {w^i_{ji}} = 1
$$

and
$$
\sum_{j = 1}^{J} {w^j_{ji}} = 1
$$

Finally, accessibility is calculated:
$$
A_i = {\sum_{j = 1}^{J} \frac {S_j{w^j_{ij}}}{\sum_{i = 1}^{n} P_i{w^i_{ij}}}}
$$

@@

```{r figure-2, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "Population in all dissemination areas in the SoBi Hamilton core service area and within 30 minutes of walking to a bike share station."}
ggplot() +
  geom_sf(data = hamilton_da_2016 %>%
            dplyr::filter(sobi_service == "Yes"),
          aes(fill = population),
          color = NA) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_fill_distiller(palette = "YlOrRd", 
                       direction = 1)
```

```{r figure-4, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "Interpolated population cells in the SoBi Hamilton core service area (outlined in black) and within 30 minutes walk to the core service area. Geographic features have been extracted. Dissemination area polygons are outlined in grey."}
ggplot() + 
  geom_sf(data = population_50x50,
          aes(color = population)) +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1) +
  geom_sf(data = hamilton_da_2016 %>%
            filter(sobi_service == "Yes"),
          fill = NA,
          color = "lightgray") + 
  geom_sf(data = sobi_service,
          fill = NA,
          colour = "blue")
```

```{r figure-5, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "Interpolated population units in the SoBi Hamilton core service area (outlined in black) and within 30 minutes walk to the core service area. Green circles represent the location of the equity stations that were added in 2018. Red circles represent the original stations. The service area of the PBSP is outlined in blue, and dissemination areas from the 2016 Canadian Census are outlined in light grey."}
ggplot() + 
  #geom_sf(fill = "antiquewhite1") +
  geom_sf(data = population_50x50, 
          aes(color = population)) +
  geom_sf(data = hamilton_da_2016 %>%
            dplyr::filter(sobi_service == "Yes"), 
          fill = NA, 
          color = "lightgray") + 
  geom_sf(data = sobi_hubs %>% 
            dplyr::filter(hub_type == "Conventional"), 
          size = 1.5, 
          shape = 21, 
          fill = "red") +
  geom_sf(data =  sobi_hubs %>% 
            dplyr::filter(hub_type == "ERI"), 
          size = 1.5, shape = 21, 
          fill = "green") +
  geom_sf(data = sobi_service, 
          fill = NA, 
          colour = "blue") +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1)
```

@@

```{r join-population-to-travel-time-data, echo=FALSE}
# current system with equity stations
ttm_walk <- left_join(ttm_walk, 
                      population_50x50 %>%
                        st_drop_geometry(),
                      by = "UID")
```

```{r join-sobi-station-to-travel-time-data, include=FALSE}
# current system with equity stations
ttm_walk <- ttm_walk %>% 
  left_join(sobi_hubs %>%
              st_drop_geometry() %>%
              dplyr::select(OBJECTID, 
                            RACKS_AMOU),
            by = c("OBJECTID" = "OBJECTID"))
```

```{r echo=FALSE}
#data frames for current and original systems
names(ttm_walk) <- c('UID', 
                     'hub', 
                     'travel_time',
                     "hub_type",
                     'population',
                     'racks')
```

```{r, echo=FALSE}
#population_sobi_original <- population_sobi_original[-c(3, 5:6)]
```

### Varying thresholds

For the system without the equity stations we remove the equity stations from the travel time matrix before calculating accessibility:
```{r}
results_3min_conventional <- ttm_walk %>%
  dplyr::filter(hub_type == "Conventional") %>%
  b2sfca(threshold = 3)
```

All stations:
```{r}
results_3min_all <- ttm_walk %>%
  b2sfca(threshold = 3)
```

Sanity check:
```{r}
sum(results_3min_conventional$los$los)
sum(results_3min_conventional$accessibility$accessibility)
```

Sanity check:
```{r}
sum(results_3min_all$los$los)
sum(results_3min_all$accessibility$accessibility)
```

For the system without the equity stations we remove the equity stations from the travel time matrix before calculating accessibility:
```{r}
results_5min_conventional <- ttm_walk %>%
  dplyr::filter(hub_type == "Conventional") %>%
  b2sfca(threshold = 5)
```

All stations:
```{r}
results_5min_all <- ttm_walk %>%
  b2sfca(threshold = 5)
```

Sanity check:
```{r}
sum(results_5min_conventional$los$los)
sum(results_5min_conventional$accessibility$accessibility)
```

Sanity check:
```{r}
sum(results_5min_all$los$los)
sum(results_5min_all$accessibility$accessibility)
```

For the system without the equity stations we remove the equity stations from the travel time matrix before calculating accessibility:
```{r}
results_10min_conventional <- ttm_walk %>%
  dplyr::filter(hub_type == "Conventional") %>%
  b2sfca(threshold = 10)
```

All stations:
```{r}
results_10min_all <- ttm_walk %>%
  b2sfca(threshold = 10)
```

Sanity check:
```{r}
sum(results_10min_conventional$los$los)
sum(results_10min_conventional$accessibility$accessibility)
```

Sanity check:
```{r}
sum(results_10min_all$los$los)
sum(results_10min_all$accessibility$accessibility)
```

For the system without the equity stations we remove the equity stations from the travel time matrix before calculating accessibility:
```{r}
results_15min_conventional <- ttm_walk %>%
  dplyr::filter(hub_type == "Conventional") %>%
  b2sfca(threshold = 15)
```

All stations:
```{r}
results_15min_all <- ttm_walk %>%
  b2sfca(threshold = 15)
```

Sanity check:
```{r}
sum(results_15min_conventional$los$los)
sum(results_15min_conventional$accessibility$accessibility)
```

Sanity check:
```{r}
sum(results_15min_all$los$los)
sum(results_15min_all$accessibility$accessibility)
```

Extract results:
```{r, include=FALSE}
los <- rbind(data.frame(results_3min_all$los, system = "All stations", threshold = "3 min"),
             data.frame(results_3min_conventional$los, system = "Conventional", threshold = "3 min"),
             data.frame(results_5min_all$los, system = "All stations", threshold = "5 min"),
             data.frame(results_5min_conventional$los, system = "Conventional", threshold = "5 min"),
             data.frame(results_10min_all$los, system = "All stations", threshold = "10 min"),
             data.frame(results_10min_conventional$los, system = "Conventional", threshold = "10 min"),
             data.frame(results_15min_all$los, system = "All stations", threshold = "15 min"),
             data.frame(results_15min_conventional$los, system = "Conventional", threshold = "15 min"))

accessibility <- rbind(data.frame(results_3min_all$accessibility, system = "All stations", threshold = "3 min"),
             data.frame(results_3min_conventional$accessibility, system = "Conventional", threshold = "3 min"),
             data.frame(results_5min_all$accessibility, system = "All stations", threshold = "5 min"),
             data.frame(results_5min_conventional$accessibility, system = "Conventional", threshold = "5 min"),
             data.frame(results_10min_all$accessibility, system = "All stations", threshold = "10 min"),
             data.frame(results_10min_conventional$accessibility, system = "Conventional", threshold = "10 min"),
             data.frame(results_15min_all$accessibility, system = "All stations", threshold = "15 min"),
             data.frame(results_15min_conventional$accessibility, system = "Conventional", threshold = "15 min"))
```

Add geometry:
```{r}
los <- los %>%
  left_join(sobi_hubs %>%
              dplyr::select(OBJECTID),
            by = c("hub" = "OBJECTID")) %>%
  st_as_sf()

accessibility <- accessibility %>%
  left_join(population_50x50 %>%
              dplyr::select(UID),
            by = c("UID" = "UID")) %>%
  st_as_sf()
```

```{r, echo=FALSE}
#accessibility plot for current system with equity stations at 3 minutes walk
ggplot() +
  geom_sf(data = population_50x50,
          color = "lightgray") +
  geom_sf(data = accessibility %>%
            dplyr::filter(threshold == "3 min"),
       aes(color = accessibility)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1) +
  facet_grid(system ~ threshold)
```

```{r, echo=FALSE}
#accessibility plot for current system with equity stations at 5 minutes walk
ggplot() +
  geom_sf(data = population_50x50,
          color = "lightgray") +
  geom_sf(data = accessibility %>%
            dplyr::filter(threshold == "5 min"),
       aes(color = accessibility)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1) +
  facet_grid(system ~ threshold)
```

```{r, echo=FALSE}
#accessibility plot for current system with equity stations at 10 minutes walk
ggplot() +
  geom_sf(data = population_50x50,
          color = "lightgray") +
  geom_sf(data = accessibility %>%
            dplyr::filter(threshold == "10 min"),
       aes(color = accessibility)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1) +
  facet_grid(system ~ threshold)
```

```{r, echo=FALSE}
#accessibility plot for current system with equity stations at 3 minutes walk
ggplot() +
  geom_sf(data = population_50x50,
          color = "lightgray") +
  geom_sf(data = accessibility %>%
            dplyr::filter(threshold == "15 min"),
       aes(color = accessibility)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1) +
  facet_grid(system ~ threshold)
```

<!--
```{r, echo=FALSE}
#level of equitable service in current system at 3 mins walk
population_sobi %>%
  mutate(Lei_min = accessibility_min - sum(sobi_los_min$sobi_los_min)/nrow(population_sobi %>% drop_na(accessibility_min))) %>%
  ggplot() +
  geom_sf(aes(color = Lei_min)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_color_gradient2()
```

@@

```{r, echo=FALSE}
ggplot() +
  geom_sf(data = population_sobi_original,
          aes(color = accessibility_o_min)) +
  geom_sf(data = hamilton_da_sobi,
          fill = NA) + 
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1)
```
-->

@@

### Threshold of 10 Minutes Walk
<!--
```{r, echo=FALSE}
dist_threshold <- 10
```

@@

```{r, echo=FALSE}
#accessibility of current system at 10 mins walk
 cowplot::ggdraw() +
 cowplot::draw_plot(map_equity, 0, 0, 1, 1) +
 cowplot::draw_plot(legend, 0.05, .6, 0.2, 0.2)
```
-->
